Bayesian exponential random graph modeling of KWL Burial network

### 2020-12-14 analysis/paper/paper.Rmd
update testing results and edit result section

<!--posterior stats for pre-E-->
We compared estimated statistics from the posterior distributions to examine their differences in  structure of the simulated networks (Table XX). For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials. The age homophily shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations. Despite positive values for some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals. Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly, indicating that physical proximity does not affect relations. For the endogenous network effects, both transitivity (gwesp) and popularity (gwdegree) present significant positive effects. The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common. The positive popularity shows a tendency toward centralization that reflects a limited number of burials have many more relations than others. This implies that a few central individuals developed personal networks through wealth accumulation in terms of the flow of goods.

<!--posterior stats for for post-E and their comparison-->
For the network after the arrival of Europeans the nodal covariates of age-homophily, sex-homophily, ritual-homophily, and the dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect. However, there are no significant effects for these nodal and dyadic covariates. We also found that the burials in the same wealth level tend not to form relations. Similar to the pre-European network, the endogenous network variables, transitivity (gwesp) and popularity (gwdegree), all demonstrate significant positive effects. The post-European network has a smaller transitivity effect and a larger popularity effect than the pre-European network. This may suggest a reduced tendency toward clustering but higher tendency toward centralization after the European presence. Both posterior estimates present symmetric distributions, with the posterior means located close to posterior medians (Table XX).

@@

<!-- comparison between two networks using bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials. To understand if this network size difference affects the posterior distributions, we conducted a sensitivity test by randomly removing `r diff_pre_post` nodes from the post-European network to resemble the size of pre-European network using a vertex bootstrap procedure. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) of a network to quantify standard errors in the network statistics of interest. This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence intervals for the network population, especially for smaller networks. We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for our two networks. The result shows that the 95% confidence interval of each network statistic contains the observed network density and degree, but the observed network clustering values are higher than the 95% CI of the bootstrap distributions. This pattern is consistent before and after node removal for the post-European network. The bootstrap distributions for the post-European network, when adjusted to be the same size pre-European network, show distinct differences from the pre-European network statistics. This indicates our networks are relatively robust to differences in sample size, and suggests a negligible sample size effect.

Similarly, we can use vertex bootstrap to compare pre- and post-European networks since it allows statistical testing for multiple networks of different sizes [@Chen2019; @Snijders1999]. We used a t-test on the distributions of network statistics from the bootstrapped network samples. The results show that there are significant differences in all network statistics from both periods. This supports the presence of anthropologically meaningful differences in the pre- and post-European networks.

@@

>>>>

<!--posterior stats for pre-E-->
We compared estimated statistics from the posterior distributions to examine their differences in structure of the simulated networks (Table XX). For the pre-European network simulations, two nodal covariates, age-homophily and ritual-homophily, have significant effects on the formation of relations between burials. The age homophily shows a negative effect that indicates burials of the same age tend to not share similar prestige goods, while the ritual-homophily demonstrates the positive effect that means burials with ritual pottery tend to form relations. Despite positive values for some covariates, such as sex-homophily and burial value-homophily, they do not show a significant tendency to form relations duo to value of zero in the confidence intervals. Similarly, the dyadic covariate, physical distance, shows positive effects but not significantly, indicating that physical proximity does not affect relations. For the endogenous network effects, transitivity (gwesp) present significant positive effect, while popularity (gwdegree) demonstrates negative effect. The positive value for transitivity suggests a tendency of burials to be clustered in closed transitive structures, indicative of the presence of multiple corporate groups sharing burial goods in common. The negative popularity shows there is a tendency toward decentralization that reflects most burials have similar number of ties without any prominent ones. This might imply that individuals have equal access to trade goods in terms of the flow of goods.

<!--posterior stats for for post-E and their comparison-->
For the network after the arrival of Europeans the nodal covariates of age-homophily, sex-homophily, ritual-homophily, and the dyadic covariate of physical distance all show positive effects, while the burial value-homophily has a negative effect. However, there are no significant effects for these nodal and dyadic covariates. We also found that the burials in the same wealth level tend not to form relations. Similar to the pre-European network, the endogenous network variable, transitivity (gwesp) demonstrates a significant positive effect. In contrast to the pre-European network, the popularity (gwdegree) shows a significant positive effect. This means there is a tendency toward centralization that reflects a limited number of burials have many more ties than others. This implies that a few central individuals may have better access to trade goods and developed personal networks through wealth accumulation. In general, the post-European network has a smaller transitivity effect and a positive popularity effect than the pre-European network. This may suggest a reduced tendency toward clustering but high tendency toward centralization after the European presence. Both posterior estimates present symmetric distributions, with the posterior means located close to posterior medians (Table XX).

@@

<!-- comparison between two networks using bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials. To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results based on Bayesian network analysis. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) of a network to quantify standard errors in the network statistics of interest [@Chen2019; @Snijders1999]. This enables the evaluation of uncertainty for networks and tests the difference between networks by examining their confidence intervals for the network population, where networks with unoverlapped intervals represent a significant difference. We set the resampling size to 1,000 networks to compute endogenous network statistics, including density, transitivity, and popularity for our two networks. The result shows a significant difference in network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach. For density and transitivity, both 95% confidence intervals show some level of overlapping that indicates no significant difference. This pattern is also similar to our results of Bayesian modeling, where both networks show similar positive or negative effects with some level of differences.

Similarly, vertex bootstrap allows statistical testing for multiple networks, such as a t-test method. We conducted a t-test on the distributions of network statistics between two networks. The results show that there are significant differences in all network statistics. This supports the presence of anthropologically meaningful differences in the pre-and post-European networks. Moreover, we tested the sample size effect by removing nodes at certain percents (5-30%) and compare their confidence intervals for both networks using the same vertex bootstrap procedure. The results show a consistent difference between the two networks that shows they are robust to the different sizes. In general, the cross-validation of results from vertex bootstrap and Bayesian modeling illustrates the robust comparison between two networks without sample size effects.

@@

Word count: `r wordcountaddin::word_count()`
